R Notebook

### 2017-06-20 manuscript/draft.Rmd

Forecasting is an emerging imperative in ecology and as such the field needs to
develop and follow good practices for conducting and evaluating ecological
forecasts. We have laid out a number of these practices (Box 1) and attempted to
implement them in a single study that builds and evaluates forecasts of
biodiversity in the form of species richness. The results of this effort are
both promising and humbling. On an absolute basis, when comparing forecasts
across sites, many different approaches to forecasting produce reasonable
forecasts. If a site is predicted to have a high number of species in the
future, relative to other sites, it generally does. However, the forecasts
perform relatively poorly at predicting variation in richness over time for any
given site. If a site is predicted to have a higher richness next year than it
did last year there is little confidence that this will be the case. This
results in simple site-level time-series, which are best predicted by simple
baselines like the long-term average and the "naive" model.

While stacked species distribution models are the most commonly used approach
for forecasting the future state of biodiversity [[cite?]], this approach
provided the worst overall forecasts of all methods evaluated. In fact, the most
common version of the stacking approach (based on rounding species' occurrence
probabilities up or down before adding them up) would have led to even worse
results than what we presented [Pellissier et al. 2013's "probabilistic
approach" paper, Calabese et al. 2014]. In general, SDM-based forecasts are
rarely evaluated, particularly in terms of repeated observations at the same
site, and especially with long time horizons. Since site-level richness is
relatively stable, the model evaluations that show high accuracy in predicting
differences among sites may not indicate much about the model's ability to
predict changes over time. This issue has also recently been recently
highlighted as an issue for using SDMs to predict the future locations of
individual species [plos one 2012 paper]. This result is particularly sobering
because this approach forms the foundation for major claims regarding the
predicted loss of biodiversity to climate change [2004 science/nature papers].
The poor performance of SDM-based forecasts highlights the crucial importance of
comparing multiple modeling approaches when conducting ecological forecasts, and
in particular the value of comparing results to simple baselines to avoid
over-interpreting the accuracy of ecological forecasts. Disciplines that have
more mature forecasting cultures often do this by reporting "forecast skill",
which is the improvement in the forecast relative to a simple baseline. We
recommend adopting this approach in future ecological forecasting research. It
also highlights the need for using hindcasting and time-series data for
assessing the effectiveness of forecasting methods, since this allows
comparisons to baselines and assessment of the predicted dynamics at each site.

The most common alternative to summed SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts []. Two
comparisons of the the SDM and direct richness modeling approaches reported that
the methods yielded equivalent results for forecasting diversity (some other
comparisons that found poor SDM performance were [[tainted]] by the rounding
error alluded to above and are not considered here). While our results generally
support rough equivalence between stacked SDMs and direct richness modeling for
point estimates, they also show that modeling richness directly
results in much better estimation of the uncertainty of the forecast and
therefore is a better overall approach to forecasting richness. A similar result
is seen when comparing joint species distribution models (JSDMs) to stacked
single species distribution models. For point estimates the joint distribution
models are roughly equivalent to stacked SDMs, but the JSDMs provide much better
estimates of uncertainty. In fact, JSDMs and modeling richness directly provide
some of the best estimates of uncertainty across all modeling approaches. This
highlights the importance evaluating models based on their uncertainty as well
as their mean estimates. Not doing so is particularly problematic in this case
because the summed species distribution models are overly confident meaning that
using them would lead to believing that richness would be restricted to a much
narrower range than would actually occur.

>>>>

Forecasting is an emerging imperative in ecology and as such the field needs to
develop and follow good practices for conducting and evaluating ecological
forecasts. We have laid out a number of these practices (Box 1) and attempted to
implement them in a single study that builds and evaluates forecasts of
biodiversity in the form of species richness. The results of this effort are
both promising and humbling. On an absolute basis, when comparing forecasts
across sites, many different approaches to forecasting produce reasonable
forecasts. If a site is predicted to have a high number of species in the
future, relative to other sites, it generally does. However, the forecasts
perform relatively poorly at predicting variation in richness over time for any
given site. If a site is predicted to have a higher richness next year than it
did last year there is little confidence that this will be the case. As a
result, the baseline models, like the long-term average and the "naive" model,
along with site-level time-series models, provided some of the best forecasts
for future biodiversity.

The most commonly used method to forecast future biodiversity, stacked SDMS
[[cite?]], provided the worst overall forecasts of all methods evaluated. In
general, SDM-based forecasts are rarely evaluated, particularly in terms of
repeated observations at the same site, and especially with long time horizons.
Since site-level richness is relatively stable, the model evaluations that show
high accuracy in predicting differences among sites may not indicate much about
the model's ability to predict changes over time. This issue has also recently
been recently highlighted as an issue for using SDMs to predict the future
locations of individual species [plos one 2012 paper]. This result is
particularly sobering because this approach forms the foundation for major
claims regarding the predicted loss of biodiversity to climate change [2004
science/nature papers]. The poor performance of SDM-based forecasts highlights
the crucial importance of comparing multiple modeling approaches when conducting
ecological forecasts, and in particular the value of comparing results to simple
baselines to avoid over-interpreting the accuracy of ecological forecasts.
Disciplines that have more mature forecasting cultures often do this by
reporting "forecast skill", which is the improvement in the forecast relative to
a simple baseline. We recommend adopting this approach in future ecological
forecasting research. It also highlights the need for using hindcasting and
time-series data for assessing the effectiveness of forecasting methods, since
this allows comparisons to baselines and assessment of the predicted dynamics at
each site.

The most common alternative to summed SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts []. Two
comparisons of the the SDM and direct richness modeling approaches reported that
the methods yielded equivalent results for forecasting diversity. While our
results generally support rough equivalence between stacked SDMs and direct
richness modeling for point estimates, they also show that modeling richness
directly results in much better estimation of the uncertainty of the forecast
and therefore is a better overall approach to forecasting richness. A similar
result is seen when comparing joint species distribution models (JSDMs) to
stacked single species distribution models. For point estimates the joint
distribution models are roughly equivalent to stacked SDMs, but the JSDMs
provide much better estimates of uncertainty. In fact, JSDMs and modeling
richness directly provide some of the best estimates of uncertainty across all
modeling approaches. This highlights the importance evaluating models based on
their uncertainty as well as their mean estimates. Not doing so is particularly
problematic in this case because the summed species distribution models are
overly confident meaning that using them would lead to believing that richness
would be restricted to a much narrower range than would actually occur.
